% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Large-scale Machine-Learning analysis of scientific PDF for following the production and the openness of research data and software in France},
  pdfkeywords={research software, research data, open access, open
science, scientometrics},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=3cm, right=3cm, top=3cm, bottom=3cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
% for compatibility with pandoc 2.10
\newenvironment{cslreferences}%
  {\setlength{\parindent}{0pt}%
  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
  {\par}

\title{Large-scale Machine-Learning analysis of scientific PDF for
following the production and the openness of research data and software
in France}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{\ldots{}}
\usepackage{authblk}
\author[%
  2%
  ]{%
  Aricia Bassinet%
  %
  %
}
\author[%
  2%
  ]{%
  Laetitia Bracco%
  %
  %
}
\author[%
  1%
  ]{%
  Anne L'Hôte%
  %
  %
}
\author[%
  1%
  ]{%
  Eric Jeangirard%
  %
  %
}
\author[%
  3%
  ]{%
  Patrice Lopez%
  %
  %
}
\author[%
  4%
  ]{%
  Laurent Romary%
  %
  %
}
\affil[1]{French Ministry of Higher Education, Research, Paris, France}
\affil[2]{University of Lorraine, France}
\affil[3]{science-miner, France}
\affil[4]{Inria, France}
\date{December 2022}

\makeatletter
\def\@maketitle{%
  \newpage \null \vskip 2em
  \begin {center}%
    \let \footnote \thanks
         {\LARGE \@title \par}%
         \vskip 1.5em%
                {\large \lineskip .5em%
                  \begin {tabular}[t]{c}%
                    \@author
                  \end {tabular}\par}%
                                                \vskip 1em{\large \@date}%
  \end {center}%
  \par
  \vskip 1.5em}
\makeatother

\begin{document}
\maketitle
\begin{abstract}
There is today no standard way for referencing research datasets and
research software in scientific communication. Emerging editorial
workflows and supporting infrastructures dedicated to dataset and
software are still poorly adopted by current publishing practices and
are highly fragmented.

To better follow the production of research datasets and software, we
present a text mining method applied to scientific publications at scale
and implemented at the French national level. Our approach relies on
state-of-the-art Machine Learning and document engineering techniques to
ensure satisfactory accuracy across multiple research areas and document
types, combining full-text harvesting, mention extraction, context
characterization and corpus-level analysis.

The annotations produced by our system are used by the French Open
Science Monitor (BSO)
\href{https://frenchopensciencemonitor.esr.gouv.fr}{platform} to follow
the production and the openness of research data and software, in the
context of the second National Plan for Open Science.

The source code and the data of the French Open Science Monitor, as well
as all the associated tools and datasets, are available under open
licences.
\end{abstract}

\textbf{Keywords}: research software, research data, open access, open
science, scientometrics

\hypertarget{introduction}{%
\section{1. Introduction}\label{introduction}}

\hypertarget{motivations}{%
\subsection{Motivations}\label{motivations}}

Datasets and software are today a core element of the research
activities, whose increasing role is broadly acknowledged {[}{]}. With
the objective of better supporting the reuse and the reproducibility of
research results, many initiatives to improve the visibility of research
dataset and software took place in the last decade, in particular
focusing on improving dataset and software cataloging {[}{]} and
standards for citation {[}{]}.

However, the usage, creation, and sharing information for the large
majority of research data and software are only available in narative
forms inside the content of scientific publications\ldots{} Among the
main reasons, we can mention the lack of incentive for researchers to
invest time on work not credited and not considered for career and
promotion, the fragmentation of policies, standards, infrastructures and
workflows, and the absence of investment by most of the scientific
publishers to implement standard for referencing of dataset and
software.

To address these pitfalls, and as a way to enforce higer standards of
openness and visibility for all research results, national Open Science
policies are currently rapidly developing\ldots{} To measure the effect
of these policies and adapt them to maximize their adoption, monitoring
tool and dashboards are crucial\ldots{}

\hypertarget{the-french-open-science-monitor}{%
\subsection{The French Open Science
Monitor}\label{the-french-open-science-monitor}}

The French Open Science Monitor, also called BSO for `Baromètre de la
Science Ouverte', is a tool for monitoring and steering the public
policy linked to the first French National Plan for Open Science (MESRI
2018). A first version of the tool was launched in 2019 providing
measurements of the rate of Open Access publications produced by all
public French research entities. A follow-up second Plan for Open
Science (MESRI 2021) has started to further promote and develop the
French open science policy, including a focus on research data and
research software. The French Open Science Monitor is updated every year
(Bracco et al. 2022), however measurements related to research datasets
and software were not covered yet.

The extension of the French Open Science Monitor to research datasets
and software was funded following a call for projects within the
framework of the French Recovery Plan. The University of Lorraine has
been asked by the Ministry of Higher Education and Research (MESR) to
lead this project alongside the MESR's Department of Decision Support
Tools and Inria.

\hypertarget{challenges}{%
\subsection{Challenges}\label{challenges}}

In view of the slow adoption of more comprehensive formal and manual
reference workflows, mining dataset and software mentions in scientific
publications appeared early as a possible solution to increase at scale
the visibility of these new key research products.

Description of prior works\ldots{} dictionary/term search, regular
expressions, domain-specific rules, restriction to XML full texts
availability. Lack of annotated data and reliable evaluations.

Characterizing the data and software usage and sharing: Often too much
``publication centric'' (e.g.~is the publication uses/shares some data
and code? versus what are the data/code product and how these data and
software are they shared and reused?)

Lack of quality metadata and considerable fragmentation of source code
repositories.

Which indicators should be selected to provide meaningful information on
data and software production and openness: how the reliability and
coverage of automatic extraction can influence these indicators? Which
minimal information should be extracted to ensure clear and trustworthy
indicators for visitors of the BSO platform?

\hypertarget{method}{%
\section{2. Method}\label{method}}

\hypertarget{machine-learning-for-mention-detection-and-characterization}{%
\subsection{2.1 Machine Learning for mention detection and
characterization}\label{machine-learning-for-mention-detection-and-characterization}}

\hypertarget{research-software-and-code}{%
\subsection{2.2 Research software and
code}\label{research-software-and-code}}

(Du et al. 2021)

\hypertarget{research-datasets}{%
\subsection{2.3 Research datasets}\label{research-datasets}}

\hypertarget{matching-and-aggregation}{%
\subsection{2.4 Matching and
aggregation}\label{matching-and-aggregation}}

\hypertarget{implementation}{%
\section{3. Implementation}\label{implementation}}

Architecture diagram

Workflow description

Infrastructure and runtime indications

\hypertarget{results}{%
\section{4. Results}\label{results}}

\hypertarget{full-text-harvesting-and-affiliation-extraction}{%
\subsection{4.1 Full text harvesting and affiliation
extraction}\label{full-text-harvesting-and-affiliation-extraction}}

\hypertarget{dataset-and-software-mention-extraction}{%
\subsection{4.2 Dataset and software mention
extraction}\label{dataset-and-software-mention-extraction}}

\hypertarget{research-product-deduplication}{%
\subsection{4.3 Research product
deduplication}\label{research-product-deduplication}}

\hypertarget{french-open-science-monitor-indicators-and-dashboards}{%
\subsection{4.4 French Open Science Monitor indicators and
dashboards}\label{french-open-science-monitor-indicators-and-dashboards}}

\ldots{}

One of the ojectives of the French Open Science Monitor is to provide
all research institutions with a local version of the indicators.
Similarly as for measuring the openness of research publications on the
existing platform, the new indicators can be bounded to datasets and
software produced by the researchers only affiliated to a particular
organization. The ability to measure the research activity at different
scale, from a given organization to the national level has led to the
creation of a BSO user group
(https://groupes.renater.fr/sympa/info/bso-etablissements).

\hypertarget{limitations-and-future-work}{%
\subsection{5 Limitations and future
work}\label{limitations-and-future-work}}

\hypertarget{limitations}{%
\subsubsection{5.1 Limitations}\label{limitations}}

\hypertarget{future-work}{%
\subsubsection{5.2 Future work}\label{future-work}}

\hypertarget{domain-coverage}{%
\paragraph{5.2.1 Domain coverage}\label{domain-coverage}}

\hypertarget{large-scale-research-entity-disambiguation}{%
\paragraph{5.2.2 Large-scale research entity
disambiguation}\label{large-scale-research-entity-disambiguation}}

\hypertarget{local-open-science-monitor}{%
\paragraph{5.2.3 Local Open Science
Monitor}\label{local-open-science-monitor}}

\hypertarget{data-and-software-availability}{%
\section{Data and software
availability}\label{data-and-software-availability}}

The data resulting from this work are publicly available on the French
Ministry of Higher Education, Research and Innovation open data portal:

The source code used for the French Open Science Monitor is available on
GitHub, and shared with open source licences.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

Thank you to the readers.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{cslreferences}
\leavevmode\hypertarget{ref-bracco_extending_2022}{}%
Bracco, Laetitia, L'Hôte, Anne, Jeangirard, Eric, and Torny, Didier.
2022. ``Extending the Open Monitoring of Open Science: A New Framework
for the French Open Science Monitor (BSO).''
\url{https://hal.archives-ouvertes.fr/hal-03651518}.

\leavevmode\hypertarget{ref-du_softcite_2021}{}%
Du, Caifan, Johanna Cohoon, Patrice Lopez, and James Howison. 2021.
``Softcite Dataset: A Dataset of Software Mentions in Biomedical and
Economic Research Publications.'' \emph{Journal of the Association for
Information Science and Technology} 72 (7): 870--84.
\url{https://doi.org/10.1002/asi.24454}.

\leavevmode\hypertarget{ref-mesri_national_2018}{}%
MESRI. 2018. ``National Plan for Open Science.''
\url{https://cache.media.enseignementsup-recherche.gouv.fr/file/Recherche/50/1/SO_A4_2018_EN_01_leger_982501.pdf}.

\leavevmode\hypertarget{ref-mesri_2nd_2021}{}%
---------. 2021. ``2nd National Plan for Open Science.''
\url{https://cache.media.enseignementsup-recherche.gouv.fr/file/science_ouverte/20/9/MEN_brochure_PNSO_web_1415209.pdf}.
\end{cslreferences}


\end{document}
